---
title: "forcats"
author: "Chun-Hui Lin"
date: "2024-11-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Some notes on data wrangling for factors from the [`forcats`](https://forcats.tidyverse.org/) site using [`starwars`](https://dplyr.tidyverse.org/reference/starwars.html) dataset as the example. 

```{r, echo=FALSE, message=FALSE}
# install.packages('forcats')
library(forcats); library(tidyverse)
```

## Add/Drop/Reverse Levels

```{r}
fct_expand(factor(c("a", "b"), levels = c("a", "b")), "c")

fct_drop(factor(c("a", "b"), levels = c("a", "b", "c")))

fct_rev(factor(c("a", "b", "c"), levels = c("a", "b", "c")))
```

## Reorder Levels

* `fct_infreq()`: order level by its frequency.

```{r}
ggplot(starwars, aes(y = fct_infreq(hair_color))) + 
  geom_bar(fill = "steelblue", width = 0.75) + 
  labs(x = "Frequency", y = "Hair color") + 
  theme_bw()
```

* `fct_inorder()`: order level by its first appearance in the dataset.

```{r}
ggplot(starwars, aes(y = fct_inorder(hair_color))) + 
  geom_bar(fill = "steelblue", width = 0.75) + 
  labs(x = "Frequency", y = "Hair color") + 
  theme_bw()
```

## Collapse Levels

* `fct_lump()`: collapse infrequent levels into one "Other" level.

  * `n=` argument: specify positive/negative n to keep most/least common n levels.
  * `prop=` argument: specify positive % to keep levels with at least % frequency.
  * `other_level=` argument: rename the lumped "Other" level.

```{r}
ggplot(starwars, aes(y = fct_infreq(fct_lump(skin_color, n = 5, other_level = "other")))) + 
  geom_bar(fill = "steelblue", width = 0.75) + 
  labs(x = "Frequency", y = "Skin color") + 
  theme_bw()
```

## Sort Levels

* `fct_reorder()`: reorder levels by another variable.

```{r}
ggplot(iris, aes(x = fct_reorder(Species, Sepal.Width), y = Sepal.Width)) +
  geom_boxplot(fill = "steelblue", width = 0.75) + 
  labs(x = "Species", y = "Sepal width") +
  theme_bw()
```
