---
title: "Power Analysis"
author: "Chun-Hui Lin"
date: "2024-12-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Some notes on power/sample size calculation from the [pwr](https://cran.r-project.org/web/packages/pwr/vignettes/pwr-vignette.html) tutorial and prior experience. Before jumping into the calculation, things-to-do are listed in the following: 

* Exposure and outcome type/distribution: determine which test to use.
* Minimal clinically important difference (MCID): expect the smallest meaningful change.  
* Power and significance level ($\alpha$): usually apply **80%** and **0.05** as the setup.
* For power calculation, number of samples collected per group is necessary. 
* For sample size calculation, effect size ($|\mu_1 - \mu_2|/\sigma$) or its components are necessary. You can either make assumption of it or obtain estimates from dataset or literature review. 

```{r, echo=FALSE}
# install.packages('pwr')
library(pwr); library(Exact)
```

## Two-sample t-test

### pwr::pwr.t.test

Assuming an effect size of 1, we’ll need at least 17 subjects in each group to detect the difference with at least 80% power using $\alpha$=0.05 and two-sided testing.

```{r}
# assume same sizes per group
pwr.t.test(d = 1, power = 0.8)
```

### pwr::pwr.t2n.test

When the sample sizes in the two groups are 198 and 37 respectively, we are able to detect a mean difference of 9.029, assuming the common standard deviation is 17.92, with at least 80% power using $\alpha$=0.05 and two-sided testing.

```{r}
# assume different sizes per group
pwr.t2n.test(n1 = 198, n2 = 37, power = 0.8)
```

## Chi-squared test

### nQuery::PTT0PU-1

`pwr.2p2n.test` is not applicable since we assume certain proportion of group 1 based on literature review. Assuming that roughly 35% of group 1 will develop the outcome, with the observed samples sizes, we are able to detect a difference in proportions of 9.3% (35% in group 1 vs. 44.3% in group 2) with 80% power using $\alpha$=0.05 and two-sided testing.

## Fisher's exact test

### Exact::power.exact.test

With an overall sample size of 30 and 5 per skin type, only a difference between 0% and 100% could be detected with 80% power using $\alpha$=0.05 and two-sided testing. Under same assumption, the differences between 0% and 80% or 20% and 100% could be detected with power between 70% and 80%. Those are large differences which may not be observed in the study with small sample size per group.

```{r}
power.exact.test(0, 1, 5, 5, method = 'fisher')
# equivalent: c(0.2, 1, 5, 5)
power.exact.test(0, 0.8, 5, 5, method = 'fisher')
```

