---
title: "Regression"
author: "Chun-Hui Lin"
date: "2024-11-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Some notes on generating regression tables from the  [tbl_regression()](https://www.danieldsjoberg.com/gtsummary/articles/tbl_regression.html) tutorial using [`trial`](https://www.danieldsjoberg.com/gtsummary/reference/trial.html) dataset as the example.

* Model choice depends on outcome type. You can modify it based on your dataset.

  * Continuous outcomes: linear regression; if the outcome is non-normally distributed, consider log-transformation. 
  * Dichotomous outcomes: logistic regression; if the outcome is very common (greater than 10%), consider poisson regression with robust error variances. 
  * Ordinal outcomes: multinomial logistic regression; if proportional odds assumption is applied, conduct [ordinal logistic regression](https://stats.oarc.ucla.edu/r/faq/ologit-coefficients/) instead.

```{r, echo=FALSE, warning=FALSE}
# install.packages('gtsummary')
library(gtsummary)
```

## gtsummary::tbl_regression()

* `tbl_regression()`

  * `exponentiate=` argument: coefficient estimates exponentiate if `TRUE`.
  * `estimate_fun=` argument: estimates round to specified digits.
  
### Linear Regression

```{r}
tbl_regression(lm(age ~ trt + grade, trial), 
               estimate_fun = ~ style_sigfig(.x, digits = 2),
               pvalue_fun = ~ style_pvalue(.x, digits = 2))
```

### Logistic Regression

```{r}
tbl_regression(glm(response ~ trt + age + grade, trial, family = 'binomial'), 
               exponentiate = T,
               estimate_fun = ~ style_ratio(.x, digits = 2),
               pvalue_fun = ~ style_pvalue(.x, digits = 2))
```

### Poisson Regression

* `coeftest()`

  * `vcov=` argument: specify sandwich covariance matrix to obtain robust standard errors.

```{r, message=F}
tidy_robust(glm(response ~ trt + age + grade, trial, family = 'poisson'), 
            exponentiate = T,
            vcov = sandwich::sandwich,
            estimate_fun = ~ style_ratio(.x, digits = 2),
            pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
  finalfit::remove_intercept() %>%
  select(term, estimate, conf.low, conf.high, p.value) %>%
  as.data.frame() %>%
  finalfit::condense_fit(estimate_name = "RR (robust SE)")
```

```{r}
# cross check using riskratio()
fit_rr = risks::riskratio(formula =response ~ trt + age + grade, approach = "robpoisson", data = trial)
broom::tidy(fit_rr, exponentiate = T)[-1, c(1:2, 6:7, 5)]
```

## Ordinal Outcome

### Multinomial Logistic Regression

```{r, echo=FALSE}
trial$trt.factor = factor(trial$trt)
```

```{r}
tbl_regression(nnet::multinom(grade ~ trt.factor, trial), 
               exponentiate = T, 
               label = trt.factor ~ 'Chemotherapy Treatment',
               estimate_fun = ~ style_ratio(.x, digits = 2),
               pvalue_fun = ~ style_pvalue(.x, digits = 2))
```

### Ordinal Logistic Regression

```{r}
broom::tidy(MASS::polr(grade ~ trt.factor, trial, Hess = T), 
            conf.int = T, exponentiate = T, p.values = T) %>%
  dplyr::filter(coef.type == 'coefficient') %>%
  select(term, estimate, conf.low, conf.high, p.value) %>%
  as.data.frame() %>%
  finalfit::condense_fit(estimate_name = "OR")
```
