---
title: "survival"
author: "Chun-Hui Lin"
date: "2024-11-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Some notes on conducting survival analyses inspiring by [Zabor's tutorial](https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html) using [`lung`](https://rdrr.io/cran/survival/man/lung.html) dataset as the example.

* Handle date-times stored as character with [`lubridate`](https://lubridate.tidyverse.org/) package.

  * Identify the order of the year (**y**), month (**m**), day (**d**), hour (**h**),
minute (**m**) and second (**s**) elements in your variable and apply the corresponding `ymd_hms()` function.

  * `%--%` operator: equivalent to `difftime()`, such as `as.duration(start.date %--% end.date)/dyears(x = 1)`.
  
* Time-to-event outcomes: if the covariate violates the proportional hazards assumption, consider adding time interaction or stratifying by itself in the model.

```{r, echo=F, message=F}
library(survival); library(survminer); library(gtsummary)
```

## Kaplan-Meier Method

### gtsummary::tbl_survfit()

#### Median Survival

```{r}
survfit(Surv(time, status == 2) ~ sex, lung) %>%
  tbl_survfit(prob = 0.5,
              label = sex ~ 'Sex',
              label_header = "**Median Survival (days)**") %>%
  add_p() %>% # equivalent: survdiff()
  modify_header(label ~ "") %>%
  bold_labels()
```

#### Median [Follow-up](https://www.bioinfo-scrounger.com/archives/follow_up_time)

The event of interest is loss-of-follow-up in this case so we utilized KM method with **reversed** status indicator. 

```{r}
survfit(Surv(time, status == 1) ~ sex, lung) %>%
  tbl_survfit(prob = 0.5,
              label = sex ~ 'Sex',
              label_header = "**Median Follow-up (days)**") %>%
  add_p() %>% # equivalent: survdiff()
  modify_header(label ~ "") %>%
  bold_labels()
```

#### 1-year Survival

```{r}
survfit(Surv(time, status) ~ sex, lung) %>%
  tbl_survfit(times = 365.25, # speify 1-year probability
              label = sex ~ 'Sex',
              label_header = "**1-year Survival (%)**", # specify column label
              estimate_fun = ~ style_percent(.x, digits = 1)) %>%
  modify_header(label ~ "") %>%
  bold_labels()
```

### survminer::ggsurvplot()

```{r, warning=F}
ggsurvplot(survfit(Surv(time, status) ~ sex, lung), 
           data = lung,
           palette = c('#E7B800', '#2E9FDF'), # customize color palettes
           linetype = 'strata', # line type by groups
           conf.int = T, # add confidence interval
           pval = T, # add p-value of log-rank test
           surv.median.line = 'hv', # add median survival info
           risk.table = T, # add risk table
           risk.table.col = 'strata', # number at risk colored by groups
           legend.title = 'Sex', # change legend title
           legend.labs = c('Male', 'Female'), # change legend labels
           xlab = 'Time in days', # customize x-axis label
           ggtheme = theme_light())
```

### survminer::pairwise_survdiff()

```{r}
# multiple comparisons of survival curves
pairwise_survdiff(Surv(ttdeath, death) ~ grade, trial)
```

## Cox PH Regression

### gtsummary::tbl_regression

```{r}
tbl_regression(coxph(Surv(ttdeath, death) ~ trt + age + grade, trial),
               exponentiate = T,
               estimate_fun = ~ style_ratio(.x, digits = 2),
               pvalue_fun = ~ style_pvalue(.x, digits = 2))
```

### survival::cox.zph()

```{r}
# test the proportional hazards assumption of model
cox.zph(coxph(Surv(ttdeath, death) ~ trt + age + grade, trial))
```

### survminer::ggforest()

* `ggforest()`

  * `refLabel=` argument: customize label for reference levels.

```{r, warning=F}
# visualize model by forest plot
ggforest(coxph(Surv(ttdeath, death) ~ trt + age + grade, as.data.frame(trial)),
         refLabel = '--')
```
