---
title: "tidyr"
author: "Chun-Hui Lin"
date: "2024-11-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Some notes on data tidying from the [`tidyr`](https://tidyr.tidyverse.org/) site.

```{r, echo=FALSE, message=FALSE}
# install.packages('tidyr')
library(tidyr); library(tidyverse)
```

```{r, echo=FALSE, message=FALSE}
daily = read_csv("output/daily.csv")
updates = read_csv("output/updates.csv")
```

## Pivoting

### Wide to Long

* `pivot_longer()`

  * `names_prefix=` argument: remove matching text form the start of each variable name.
  * `names_sep=` argument: specify positions or regular expression to break on.
  * `names_pattern=` argument: specify regular expression to split variable name into multiple columns.
  * `names_transform=` argument: convert variable type to integer. 
  * `values_drop_na=` argument: drop `NA` rows if `TRUE`.

* Regular Expressions

  * `.`: matches any character. 
  * `?`: 0 or 1.
  * `*`: 0 or more.

```{r}
# Numeric data in columns names
billboard %>% 
  pivot_longer(
    cols = tidyselect::starts_with("wk"), 
    names_to = "week", 
    names_prefix = "wk",
    names_transform = as.integer, # equivalent: readr::parse_number
    values_to = "rank",
    values_drop_na = TRUE,
  )
```

```{r}
# Many variables in column names
(who_wide = who %>% 
  pivot_longer(
    cols = new_sp_m014:newrel_f65,
    names_to = c("diagnosis", "gender", "age"), 
    names_pattern = "new_?(.*)_(.)(.*)",
    names_transform = list(
      gender = as.factor, # equivalent: readr::parse_factor
      age = ~ readr::parse_factor(
        .x,
        levels = c("014", "1524", "2534", "3544", "4554", "5564", "65"), 
        ordered = TRUE
      )
    ),
    values_to = "count",
    values_drop_na = TRUE,
  ))
```

```{r}
# Multiple observations per row
household %>% 
  pivot_longer(
    cols = !family, 
    names_to = c(".value", "child"), # part of the column name specify the value being measured
    names_sep = "_", 
    values_drop_na = TRUE
  )
```

### Long to Wide

* `pivot_wider()`

  * `name_glue=` argument: use `name_from` columns to create custom column names.
  * `name_expand=` argument: show implicit factor levels if `TRUE`.
  * `values_fill=` argument: specify the value filled in with when missing.
  * `values_fn=` argument: specify the function apply to the value.
  * `unused_fn=` argument: summarize the values from the unused column by specified function.

```{r}
# Capture-recapture data
fish_encounters %>% 
  pivot_wider(
    names_from = station, 
    values_from = seen,
    values_fill = 0
  )
```

```{r}
# Aggregation
warpbreaks %>% 
  pivot_wider(
    names_from = wool, 
    values_from = breaks,
    values_fn = mean
  )
```

```{r}
# Generate column name from multiple variables
who_wide %>%
  pivot_wider(
    names_from = c(diagnosis, gender, age),
    values_from = count,
    names_glue = "cnt_{diagnosis}_{gender}_{age}"
  )
```

```{r}
# Tidy census
daily %>% 
  pivot_wider(
    names_from = day, 
    values_from = value, 
    names_expand = TRUE,
    values_fill = 0
  )
```

```{r}
# Unused columns
updates %>% 
  pivot_wider(
    id_cols = county, 
    names_from = system, 
    values_from = value,
    unused_fn = list(date = max)
  )
```
